---
title: "Exam 1"
---

#Grand Research Question: What does the consumption of each food category in each country look like?


**Get Data**
```{r}
library(tidytuesdayR)
library(tidyverse)
library(sf)
library(rnaturalearth)

tuesdata <- tt_load('2020-02-18')
head(tuesdata)

fc <- tuesdata$food_consumption
head(fc)


```

**Understand Data**
```{r}
nrow(fc)
head(fc)
str(fc)
dim(fc)
tail(fc)

```

**Understand Variables Individually**

food_category is a categorical variable so use a bar graph to visualize
consumption is quantitative variable, could use histogram, density, or boxplot to visualize
co2_emmission is  quantitative variable, could use histogram, density, or boxplot to visualize
country is categorical variable so could use a bar graph to visualize


##Answering Grand research question


```{r}
library(ggplot2)

fc1 <- fc|>
  group_by(country, food_category) |>
  summarise(mean = mean(consumption)) 

```

```{r}
df_beef <- fc1 |>
  filter(food_category == "Beef")

world_boundaries <- ne_countries(returnclass = "sf")
ggplot(world_boundaries) + 
  geom_sf(data =df_beef,  aes(fill = mean))

library(maps)

world <- map_data("world")

ggplot(world, aes(map_id = region)) +
  geom

```




