---
title: "Exam 1"
Author: "Cristian Cano"
Date: 05/06/2025
---

```{r, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
```


# Grand Research Question: What does the consumption of each food category in each country look like?


**Get Data**
```{r}
library(tidytuesdayR)
library(tidyverse)
library(sf)
library(rnaturalearth)
library(maps)

tuesdata <- tt_load('2020-02-18')
head(tuesdata)

fc <- tuesdata$food_consumption
head(fc)

```


**Understand Data**
```{r}
nrow(fc)
head(fc)
str(fc)
dim(fc)
tail(fc)
fc
```

**Understand Variables Individually**

- food_category is a categorical variable so use a bar graph to visualize
- consumption is quantitative variable, could use histogram, density, or boxplot to visualize
- co2_emmission is  quantitative variable, could use histogram, density, or boxplot to visualize
- country is categorical variable so could use a bar graph to visualize



## Answering grand research question


```{r}
library(ggplot2)

#Making a new dataset grouping by country and food category
fc1 <- fc|>
  group_by(country, food_category)

#view(fc1)

```

```{r}
#Filtering for only the food category Beef
df_beef <- fc1 |>
  filter(food_category == "Beef")
head(df_beef)

#Making a world map
world_map <- map("world", plot = FALSE, fill = TRUE)
world_sf <- st_as_sf(world_map)

#Joining world map to country
joined_world_sf_beef <- world_sf |>
  left_join(df_beef, by = c("ID" = "country"))

#Mapping consumption of the food category
joined_world_sf_beef |>
  ggplot() +
  geom_sf(aes(fill = consumption), na.rm = TRUE) +
  scale_fill_gradient(
    low  = "lightyellow",
    high = "darkred",
    name = "Consumption (kg/person/year)"
  ) +
  labs(title = "Consumption of Beef by Country") +
  theme_minimal()

```
- There is alot of Beef consumption in Argentina, compared to the rest of the world.
```{r}
df_eggs <- fc1 |>
  filter(food_category == "Eggs")
head(df_eggs)

joined_world_sf_eggs <- world_sf |>
  left_join(df_eggs, by = c("ID" = "country"))

joined_world_sf_eggs |>
  ggplot() +
  geom_sf(aes(fill = consumption), na.rm = TRUE) +
  scale_fill_gradient(
    low  = "lightyellow",
    high = "darkred",
    name = "Consumption (kg/person/year)"
  ) +
  labs(title = "Consumption of Eggs by Country") +
  theme_minimal()
```


```{r}
df_fish <- fc1 |>
  filter(food_category == "Fish")

joined_world_sf_fish <- world_sf |>
  left_join(df_fish, by = c("ID" = "country"))

joined_world_sf_fish |>
  ggplot() +
  geom_sf(aes(fill = consumption), na.rm = TRUE) +
  scale_fill_gradient(
    low  = "lightyellow",
    high = "darkred",
    name = "Consumption (kg/person/year)"
  ) +
  labs(title = "Consumption of Fish by Country") +
  theme_minimal()
```

```{r}
df_lamb <- fc1 |>
  filter(food_category == "Lamb & Goat")

joined_world_sf_lamb <- world_sf |>
  left_join(df_lamb, by = c("ID" = "country"))

joined_world_sf_lamb |>
  ggplot() +
  geom_sf(aes(fill = consumption), na.rm = TRUE) +
  scale_fill_gradient(
    low  = "lightyellow",
    high = "darkred",
    name = "Consumption (kg/person/year)"
  ) +
  labs(title = "Consumption of Lamb & Goat by Country") +
  theme_minimal()
```

```{r}
df_milk <- fc1 |>
  filter(food_category == "Milk - inc. cheese")

joined_world_sf_milk <- world_sf |>
  left_join(df_milk, by = c("ID" = "country"))

joined_world_sf_milk |>
  ggplot() +
  geom_sf(aes(fill = consumption), na.rm = TRUE) +
  scale_fill_gradient(
    low  = "lightyellow",
    high = "darkred",
    name = "Consumption (kg/person/year)"
  ) +
  labs(title = "Consumption of Milk & Cheese by Country") +
  theme_minimal()
```

```{r}
df_nuts <- fc1 |>
  filter(food_category == "Nuts inc. Peanut Butter")

joined_world_sf_nuts <- world_sf |>
  left_join(df_nuts, by = c("ID" = "country"))

joined_world_sf_nuts |>
  ggplot() +
  geom_sf(aes(fill = consumption), na.rm = TRUE) +
  scale_fill_gradient(
    low  = "lightyellow",
    high = "darkred",
    name = "Consumption (kg/person/year)"
  ) +
  labs(title = "Consumption of Nuts (including peaunut butter) by Country") +
  theme_minimal()
```

```{r}
df_pork <- fc1 |>
  filter(food_category == "Pork")

joined_world_sf_pork <- world_sf |>
  left_join(df_pork, by = c("ID" = "country"))

joined_world_sf_pork |>
  ggplot() +
  geom_sf(aes(fill = consumption), na.rm = TRUE) +
  scale_fill_gradient(
    low  = "lightyellow",
    high = "darkred",
    name = "Consumption (kg/person/year)"
  ) +
  labs(title = "Consumption of Pork by Country") +
  theme_minimal()
```

```{r}
df_poultry <- fc1 |>
  filter(food_category == "Poultry")

joined_world_sf_poultry <- world_sf |>
  left_join(df_poultry, by = c("ID" = "country"))

joined_world_sf_poultry |>
  ggplot() +
  geom_sf(aes(fill = consumption), na.rm = TRUE) +
  scale_fill_gradient(
    low  = "lightyellow",
    high = "darkred",
    name = "Consumption (kg/person/year)"
  ) +
  labs(title = "Consumption of Poultry by Country") +
  theme_minimal()
```

```{r}
df_rice <- fc1 |>
  filter(food_category == "Rice")

joined_world_sf_rice <- world_sf |>
  left_join(df_rice, by = c("ID" = "country"))

joined_world_sf_rice |>
  ggplot() +
  geom_sf(aes(fill = consumption), na.rm = TRUE) +
  scale_fill_gradient(
    low  = "lightyellow",
    high = "darkred",
    name = "Consumption (kg/person/year)"
  ) +
  labs(title = "Consumption of Rice by Country") +
  theme_minimal()
```
- Rice consumption is concentrated mainly in Asia

```{r}
df_soybeans <- fc1 |>
  filter(food_category == "Soybeans")

joined_world_sf_soybeans <- world_sf |>
  left_join(df_soybeans, by = c("ID" = "country"))

joined_world_sf_soybeans |>
  ggplot() +
  geom_sf(aes(fill = consumption), na.rm = TRUE) +
  scale_fill_gradient(
    low  = "lightyellow",
    high = "darkred",
    name = "Consumption (kg/person/year)"
  ) +
  labs(title = "Consumption of Soybeans by Country") +
  theme_minimal()
```
- This is alot of soybean consumption in Zambia, South Korea, and Japan, less in other parts of the world.

```{r fig.alt = "Consumption of Wheat by Country"}


df_wheat <- fc1 |>
  filter(food_category == "Wheat and Wheat Products")

joined_world_sf_wheat <- world_sf |>
  left_join(df_wheat, by = c("ID" = "country"))

joined_world_sf_wheat |>
  ggplot() +
  geom_sf(aes(fill = consumption), na.rm = TRUE) +
  scale_fill_gradient(
    low  = "lightyellow",
    high = "darkred",
    name = "Consumption (kg/person/year)"
  ) +
  labs(title = "Consumption of Wheat by Country") +
  theme_minimal()
```

- There is a lot of Wheat consumption in North Africa



